---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# USGSlidar

<!-- badges: start -->
<!-- badges: end -->

The USGSlidar package was developed to make it somewhat easy to query the collection of
lidar data available from the U.S. Geological Survey's 3D Elevation Program (and other sources)
to find out if data are available for specific locations of interest. The package 
accesses some of the project and tile index files available from 3DEP and uses them to
search for data. The package also provides links to lidar data files that can be downloaded
for free directly from 3DEP servers.

The package also facilitates access to the USGS lidar collection maintained in Entwine format.
This collection contains a subset of the data available from USGS but it is constantly being
updated to match the data available in the national map. The index for the Entwine collection
does not contain detailed project information so there is extra effort required to marry the
Entwine index with information in the USGS 3DEP collection.

**This code should be considered experimental. The examples work but some options for functions
may not work as expected. I have used the code for several months but I tend to use it the same
way all the time so I haven't fully exercised all of the options for the functions. At some point,
I will add more tests and clean unp things that don't work.**

## Installation

This package is only distributed from my GitHub account. It may make it to CRAN at some point but
no promimses.

~~You can install the released version of USGSlidar from [CRAN](https://CRAN.R-project.org) with:~~

``` r
#install.packages("USGSlidar")
```

**USGSlidar** is currently available as a development version only. The **devtools** package is required to
install **USGSlidar**. If you have not previously used **devtools**, use the commented line of
code to install the package. Note that this will also install several additional packages needed
for devtools. If you do not want the vignettes, set *build_vignettes = FALSE*.

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
library(devtools)
devtools::install_github("bmcgaughey1/USGSlidar", build_vignettes = TRUE)
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(USGSlidar)
## basic example code
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

