---
title: "USGSlidar -- Basics"
author: "Robert J. McGaughey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{USGSlidar -- Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overall Goal

The USGSlidar package was developed to make it somewhat easy to query the collection of
lidar data available from the U.S. Geological Survey's 3D Elevation Program (and other sources)
to find out if data are available for specific locations of interest. The package 
accesses some of the project and tile index files available from 3DEP and uses them to
search for data. The package also provides links to lidar data files that can be downloaded
for free directly from 3DEP servers.

The package also facilitates access to the USGS lidar collection maintained in Entwine format.
This collection contains a subset of the data available from USGS but it is constantly being
updated to match the data available in the national map. The index for the Entwine collection
does not contain detailed project information so there is extra effort required to marry the
Entwine index with information in the USGS 3DEP collection.

## Packages

These are some libraries that are useful when working with the USGSlidar package. The 
following packages are loaded with the USGSlidar package:
* jsonlite
* PROJ
* raster
* RCurl
* rgdal
* rgeos
* sp
* sf
* stats
* urltools

These packages are used for this vignette:
* mapview
* RSQLite
* dplyr
* tidyverse
* lubridate


```{r setup, echo=T, results='hide', error=F, warning=F, message=F}
library(rgdal)
library(sf)
library(PROJ)
library(raster)
library(rgeos)
library(mapview)
library(RSQLite)
library(dplyr)
library(tidyverse)
library(lubridate)
library(USGSlidar)
```

